---
title: "Doublet Detection on 8k PBMCs from 10X Genomics"
author: "Tom Kelly and Jonathon Shor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Install packages

### Install package:

```{r}
if(!require(devtools)){
  install.packages("devtools") # If not already installed
}
devtools::install_github("TomKellyGenetics/DoubletDetection", ref = "r-implementation")
```

### Install dependancies:


```{r, include = FALSE, echo = FALSE}
options(repos = "https://cran.stat.auckland.ac.nz/")
```



```{r, eval = FALSE}
#devtools::install_github("JinmiaoChenLab/Rphenograph")
devtools::install_github("TomKellyGenetics/Rphenograph")
install.packages("hdf5r")
install.packages("Matrix")
```

### Install hdf5 (if needed)

OS X (using Homebrew)	`brew install hdf5`

Debian-based systems (including Ubuntu)	`sudo apt-get install libhdf5-dev`

Systems supporting yum and RPMs	`sudo yum install hdf5-devel`

HDF5 1.8.14 has been pre-compiled for Windows and is available at https://github.com/mannau/h5-libwin 

### load package (and dependancies)

```{r}
library("DoubletDetection")
```

## Download Data from 10X Genomics

These commands will be called from R in Linus and Mac systems. The files will need to be downloaded from the given url (into the working directory) in Windows.

```{r}
#download files
if(!(file.exists("pbmc8k_filtered_gene_bc_matrices.tar.gz"))){
  system("wget http://cf.10xgenomics.com/samples/cell-exp/2.1.0/pbmc8k/pbmc8k_filtered_gene_bc_matrices.tar.gz")
  #extract compressed files 
  #system("tar -xvzf pbmc8k_filtered_gene_bc_matrices.tar.gz")
}
```

## Load Counts Matrix

```{r}
matrix_path <- 'filtered_gene_bc_matrices/GRCh38/matrix.mtx'
raw_counts <- DoubletDetection::load_mtx(matrix_path)
# Remove columns with all 0s
non_zero_genes <- apply(raw_counts, 1, function(x) sum(x) != 0)
raw_counts <- raw_counts[non_zero_genes, ]
```

## Run Doublet Detection

Right now, phenograph is a bit talkative, so we suppress the output to avoid lots of text:

```{r, include = FALSE}
#initialise BoostClassifier Reference Class (with analysis parameters)
clf <- DoubletDetection::BoostClassifier$new(n_iters=50)
 
#run doublet prediction
start <- Sys.time()
doublets <- clf$fit(raw_counts)$predict()
end <- Sys.time()
```

```{r}
print(paste("Time elapsed", end-start,"seconds,", (end-start) / clf.n_iters, clf.n_iters), "sec/iteration, for", n_inters, "iterations")
```

## Visualize Results

### Convergence of Doublet Calls

```{r, eval = FALSE}
f <- DoubletDetection::plot$convergence(clf, save='convergence_test.pdf', show=TRUE)
```


### Doublets on tSNE map

```{r, eval = FALSE}
f2_tsne_coords <- DoubletDetection::plot$tsne(raw_counts, doublets, save='tsne_test.pdf', show=True)
```


```{r, echo = FALSE, include = FALSE}
save.image("vignette.RData")
```
